---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
 
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# ymlthis: write YAML for R Markdown, bookdown, blogdown, and more

<!-- badges: start -->
[![Build Status](https://travis-ci.com/rstudio-education/ymlthis.svg?branch=master)](https://travis-ci.com/rstudio-education/ymlthis)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/lwqfvb40js0u01fs/branch/master?svg=true)](https://ci.appveyor.com/project/malcolmbarrett/ymlthis-n0nhc/branch/master)
[![Codecov test coverage](https://codecov.io/gh/rstudio-education/ymlthis/branch/master/graph/badge.svg)](https://codecov.io/gh/rstudio-education/ymlthis?branch=master)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/ymlthis)](https://cran.r-project.org/package=ymlthis)
<!-- badges: end -->

ymlthis makes it easy to write YAML front matter for R Markdown and related documents. `yml_*()` functions write functions and `use_*()` functions let you write the resulting YAML to your clipboard or to `.yaml` files related to your project.

## Installation

You can install the development version of ymlthis from GitHub with:

``` r
# install.packages("remotes")
remotes::install_github("rstudio-education/ymlthis")
```

## Example

`yml()` creates a basic `yml` object returns simple YAML with the author and date.

```{r example}
library(ymlthis)

yml()
```

ymlthis supports many YAML arguments, with YAML-generating functions prefixed with `yml_*()`:

```{r example-2}
yml() %>% 
  yml_author(c("Yihui Xie", "Hadley Wickham"), affiliation = "RStudio") %>% 
  yml_date(lubridate::today()) %>% 
  yml_output(
    pdf_document(keep_tex = TRUE, includes = includes2(after_body = "footer.tex")), 
    bookdown::html_document2()
  ) %>% 
  yml_latex_opts(biblio_style = "apa-like")
```
